Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-06 measurements.tex
Reorganizing the document and adding basic conclusions

\subsection{Goodput}
For the goodput, we separate the results according to their time benchmarks
as we did in Section \ref{sec:metrics} sicne their consider different
methodologies. We proceed first with the measurements related to the encoder
and later review the case of the decoder.

Fig. \ref{fig:enc_good_rasp1} shows the results of the encoder goodput
measurements for the \ac{Raspi} 1. Figs. \ref{fig:enc_good_rasp1_gen_gf2}
and \ref{fig:enc_good_rasp1_gen_gf256} present goodput as function of the
generation size for $GF(2)$ and $GF(2^8)$ respectively for a packet size
of $B = 1600$ bytes. Figs. \ref{fig:enc_good_rasp1_packet_gf2} and
\ref{fig:enc_good_rasp1_packet_gf256} present the same metric but now as
function of the packet size for $GF(2)$ and $GF(2^8)$ respectively. In this
case the generation size was set to $g = 16$ packets.

@@

For the plots in the mentioned figures from the encoder goodput measurements,
\ac{RLNC} presents the worst performance in terms of goodput and
\ac{TSNC} with $dep = 16$, the best regardless of the \ac{Raspi} model.
Given that the processing time depends in the amount of coded packets
required to create, \ac{RLNC} is the slowest to process since it must
use all of the $g$ original packets. Later, sparse codes process the data
at a larger rate since less packets are being mixed when creating a
coded packet. The caveat of these schemes is that the sparser the code,
the more probable the ocurrences of \ac{l.d.} packets is. So, basically
the sparser the codes, then more overhead due to transmissions of \ac{l.d.}
packets they have since it is more likely to generate the same set of
packets. Excluding the coding coefficients overhead, the overhead due to
transmissions of \ac{l.d.} packets can be very high.
For example, if we consider \ac{TSNC} with $dep=16$ and $g = 16$, the budget
in this case permits to send up to 32 packets which is $2\times$ the generation
size for an overhead of $100\%$ excluding the one from the coding
coefficients. This happens because \ac{TSNC} is allowed to add too much
redundancy in this case, which permits it to use a very low code density $d$.
For \ac{RLNC} this is not the case, since the occurrence of \ac{l.d.} coded
packets is low. Even for $GF(2)$, the average amount of redundancy has proven
to be 1.6 packets after $g$ have been transmitted, much less than the
cases where sparse codes are employed. Overall, we observe that there is
a trade-off between goodput and \ac{l.d.} coded packets transmission
overhead.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/23_07_2015/goodput_vs_generation_size_Rasp_v2_encoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:enc_good_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/23_07_2015/goodput_vs_generation_size_Rasp_v2_encoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:enc_good_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/23_07_2015/goodput_vs_symbol_size_Rasp_v2_encoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/23_07_2015/goodput_vs_symbol_size_Rasp_v2_encoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder goodput measurements for the \ac{Raspi} 2}
    \label{fig:enc_good_rasp2}
\end{figure*}

Figures description

>>>>

\subsection{Goodput}
For the goodput, we separate the results according to their time benchmarks
as we did in Section \ref{sec:metrics} since their consider different
methodologies. We proceed first with the measurements related to the encoder
and later review the case of the decoder.

\subsubsection{Encoding}
Fig. \ref{fig:enc_good_rasp1} shows the results of the encoder goodput
measurements for the \ac{Raspi} 1. Figs. \ref{fig:enc_good_rasp1_gen_gf2}
and \ref{fig:enc_good_rasp1_gen_gf256} present goodput as function of the
generation size for $GF(2)$ and $GF(2^8)$ respectively for a packet size
of $B = 1600$ bytes. Figs. \ref{fig:enc_good_rasp1_packet_gf2} and
\ref{fig:enc_good_rasp1_packet_gf256} present the same metric but now as
function of the packet size for $GF(2)$ and $GF(2^8)$ respectively. In this
case the generation size was set to $g = 16$ packets.

@@

For the plots in the mentioned figures from the encoder goodput measurements,
\ac{RLNC} presents the worst performance in terms of goodput and
\ac{TSNC} with $dep = 16$, the best regardless of the \ac{Raspi} model.
Given that the processing time depends in the amount of coded packets
required to create, \ac{RLNC} is the slowest to process since it must
use all of the $g$ original packets. Later, sparse codes process the data
at a larger rate since less packets are being mixed when creating a
coded packet. The caveat of these schemes is that the sparser the code,
the more probable the ocurrences of \ac{l.d.} packets is. So, basically
the sparser the codes, then more overhead due to transmissions of \ac{l.d.}
packets they have since it is more likely to generate the same set of
packets. Excluding the coding coefficients overhead, the overhead due to
transmissions of \ac{l.d.} packets can be very high.
For example, if we consider \ac{TSNC} with $dep=16$ and $g = 16$, the budget
in this case permits to send up to 32 packets which is $2\times$ the generation
size for an overhead of $100\%$ excluding the one from the coding
coefficients. This happens because \ac{TSNC} is allowed to add too much
redundancy in this case, which permits it to use a, single in this case,
very low code density $d$. For \ac{RLNC} this is not the case, since the
occurrence of \ac{l.d.} coded packets is low. Even for $GF(2)$, the
average amount of redundancy has proven to be 1.6 packets after $g$ have
been transmitted, much less than the cases where sparse codes are
utilized. Overall, we observe that there is a trade-off between goodput
and \ac{l.d.} coded packets transmission overhead.

For Figs.~\ref{fig:enc_good_rasp1_packet_gf2},
\ref{fig:enc_good_rasp1_packet_gf256}, \ref{fig:enc_good_rasp2_packet_gf2}
and \ref{fig:enc_good_rasp2_packet_gf256}; we see the packet
size effect in the encoding goodput in both \ac{Raspi} models and for
a fixed generation size. For \ac{TSNC} in this case, we allow for 5 density
changes during the generation transmission and again consider the same
budgets as before. In all the cases, we clearly observe that there is an
optimal packet size to operate with. When reviewing the specifications
of the \ac{Raspi}, both models use a L1 cache of 16 KiB. Hence, the trends
in the packet size can be explained as follows: For low packet sizes,
the data processing does not overload the \ac{CPU} of the \ac{Raspi}, so
the goodput progressively increases given that we simply process more
data as the packet size increases. However, after a certain packet size,
the goodput gets affected since the cache starts to saturate. The packets
towards the cache needs to have more processing. Also important, the
\ac{CPU} needs to allocate computational resources to other processes from
the different tasks running in the \ac{Raspi}. Beyond this critical packet
size, the \ac{CPU} just queues processing which incurs in larger delay
reducing the goodput.

If we consider the previous effect when reviewing the trends in the
mentioned figures,
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/23_07_2015/goodput_vs_generation_size_Rasp_v2_encoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:enc_good_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/23_07_2015/goodput_vs_generation_size_Rasp_v2_encoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:enc_good_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/23_07_2015/goodput_vs_symbol_size_Rasp_v2_encoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/23_07_2015/goodput_vs_symbol_size_Rasp_v2_encoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:enc_good_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Encoder goodput measurements for the \ac{Raspi} 2}
    \label{fig:enc_good_rasp2}
\end{figure*}
%
\subsubsection{Decoding}
